---
title: "Loops and Functions Activity"
author: "Lewis White"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Task 1

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thurday", "Friday", "Saturday", "Sunday")
transects <- c("Transect A", "Transect B", "Transect C")
```

```{r}
week_matrix <- matrix(nrow = length(weekdays), ncol = length(transects))

for (i in seq_along(transects)) {
  for (j in seq_along(weekdays)) {
    day_trans <- paste(weekdays[j], "-", transects[i])
    week_matrix[j, i] <- day_trans
  }
}
```

## Task 2

Write a function called force that calculates a force (in Newtons), given inputs of mass (in kg) and acceleration (in ms2 (recall: F=ma), and returns a statement “The resulting force is ___ Newtons.”


```{r}
force <- function(mass, acceleration) {
  f = mass * acceleration
  print(paste("The resulting force is", f, "Newtons"))
}

force(mass = 2, acceleration = 3)
```


## Task 3
The length:weight relationship for fish is: W=aLb, where where L is total fish length (centimeters), W is the expected fish weight (grams), and a and b are species-dependent parameter values (shown below for several fish from Peyton et al. 2016).

Recreate the table above as a data frame stored as fish_parms. Then, write a function called fish_weight that allows a user to only enter the common name (argument fish_name) and total length (argument tot_length) (in centimeters) of a fish, to return the expected fish weight in grams. Test it out for different species and lengths.

Now, try creating a vector of lengths (e.g. 0 to 100, by increments of 1) and ensuring that your function will calculate the fish weight over a range of lengths for the given species (try this for milkfish, storing the output weights as milkfish_weights.

```{r}
fish_parms <- tribble(
  ~sci_name, ~common_name, ~a_est, ~b_est,
  "Chanos chanos", "Milkfish", 0.0905, 2.52,
  "Sphyraena barracuda", "Great barracuda", 0.0181, 3.27,
  "Caranx ignobilis", "Giant trevally", 0.0353, 3.05)
  
fish_weight <- function (fish_name, tot_length) {
  fish_subset <- filter(.data = fish_parms, common_name == fish_name)
  weight = fish_subset$a_est * tot_length * fish_subset$b_est
  print(weight)
}

fish_weight(fish_name = "Milkfish", tot_length =  12)


fish_lengths <- seq(from = 0, to = 100, by = 1)

milkfish_weights <- fish_weight(fish_name = "Milkfish", tot_length = fish_lengths)
```


## Task 4


Wave power (more accurately wave energy flux) in deep water is approximated by:

Pdeep=0.5H2T
where P is power in kWm (potential power per wave meter), H is wave height in meters (more specifically, the significant wave height), and T is the wave period in seconds. Learn more here.

Write a function called wave_power that calculates potential ocean wave power given inputs of wave height and period.

Use your wave_power function to approximate wave power for a period of 8 seconds, over a range of wave heights from 0 to 3 meters by increments of 0.2 meters.


```{r}
wave_power <- function(height, period) {
  power = 0.5 * height^2 * period
  print(power)
}

heights <- seq(from = 0, to = 3, by = 0.2)

wave_power(height = heights, period = 8)

wave_power(height = seq(from = 0, to = 3, by = 0.2), period = 8)
```


## Task 5

the wave energy equation changes based on ocean depth. Along the coast of Brenville, which has a very sharp shelf as the wave approaches the coast, wave energy is approximated using the deep ocean equation (the one you used above) for depths > 12 meters, and a shallow equation for depths <= 12 meters. The Brenville team estimates shallow wave power by:

Pshallow = 0.81H2T

Create a function that requires inputs of water depth, wave height and period, then returns the approximated wave power using the correct equation for the depth entered. It should also include a message (hint: use message() just like you would use warning!) that lets the user know if the shallow or deep water equation was used.


```{r}
wave_power <- function(depth, height, period) {
  
  if (depth > 12) {
    print("deep water equation used")
    power = 0.5 * height^2 * period
    print(power)
  }
  
  if (depth <= 12) {
    print("shallow water equation used")
    power = 0.81 * height^2 * period
    print(power)
  }
}

wave_power(13, 12, 8)
```

